Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2021-12-17 ms/response_JEB_20211024.txt

REVIEWER1. Among the differences that can be found on the three studied habitats, authors mostly talk about differences in light incidence. Are there any differences also in background colour? Predator community? Anole community? If so, how will they contribute or not to the obtained results?

(R) Background color definitely varies between habitats, which can be seen with the naked eye from photos, but we did not quantify this precisely (we do not have irradiance profiles for the habitats telling us which wavelengths are different). Anole and predator communities have been shown to vary between islands (at least in density of the main species, which are overall present across all islands), but we do not know whether they also vary between the habitats within islands.

REVIEWER1. How big is the home range of an Anolis? Sorry of I missed it. This information will help understanding the importance of your experimental design

(R) Ambika's paper reports wider territories than previously thought in A. sagrei: 225m2 for males (which is 15 by 15m), and 36m2 for females (6 by 6m), with with evidence for multiple matings in both sexes. These estimates were measured on a campus in Florida and may not represent the real territory size on our study islands, however. Nevertheless, although the species is still classified as territorial, its ability for movement and promiscuity, its continuous distribution at high densities throughout the islands and the polyphyly of individuals from different habitats within islands suggests that long-range gene flow is highly plausible. (Do you guys know something about the cruising range of an anolis...)

@@

REVIEWER1. In line 91, what do you mean by climatic variables? Light incidence?

(R) We have now added that these climatic variables include precipitations and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.

### 2021-12-17 ms/sections/abstract.tex
implemented Graham's comments on the pdf

Animal signals evolve in an ecological context. Locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation. Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales. Here, we investigated differentiation of dewlap coloration among habitat types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales. We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands, namely beach scrub, primary coppice forest and mangrove forest. Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances. Passive divergence in dewlap phenotype associated with isolation-by-distance did not \textcolor{olive}{seem to} explain our results. On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported. We suggest that neutral processes or selection could be involved in several ways, including sexual selection. Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

>>>>

Animal signals evolve in an ecological context. Locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation. Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales. Here, we investigated differentiation of dewlap coloration among habitat types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales. We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands, namely beach scrub, primary coppice forest and mangrove forest. Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances. Passive divergence in dewlap phenotype associated with isolation-by-distance did not \textcolor{olive}{seem to} explain our results. On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported. We suggest that neutral processes or selection could be involved in several ways, including sexual selection. Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving \sout{these} patterns \textcolor{olive}{of dewlap color variation across islands}.

### 2021-12-17 ms/sections/discussion.tex
implemented Graham's comments on the pdf

\textcolor{olive}{Genetic drift may contribute to some of the observed variation. Indeed, while only Abaco showed significant patterns consistent with isolation-by-distance (which may emerge under limited dispersal and drift, \citealt{Wright1943, Kimura1964, Slatkin1987}), there may have been too few sites on most islands to conclusively detect it, and even then, the absence of detectable isolation-by-distance may not necessarily constitute evidence for the absence of drift. Besides, spatial autocorrelation was the strongest on islands sampled at the largest scales (e.g. some sites on Abaco were nearly 100km apart, and Eleuthera -- the second strongest signal, albeit nonsignificant -- had sites more than 30km apart), such that it is possible that neutral processes and/or dispersal limitations might in fact contribute to shaping variation over long distances. That said, many significant differences were found between habitats in close proximity, contrary to what would be expected under isolation-by-distance, including on islands where spatial autocorrelation was detected. Moreover, \textit{A. sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations particularly prone to drift. Together with the fact that isolation-by-distance may not necessarily only emerge from drift (e.g. if there is a spatial environmental gradient), this indicates that genetic drift may have limited potential to explain the differences observed between habitats, at least at a local scale.}\\

>>>>

\textcolor{olive}{Genetic drift could contribute to some of the observed variation. Indeed, while only Abaco showed significant patterns consistent with isolation-by-distance (which may emerge under limited dispersal and drift, \citealt{Wright1943, Kimura1964, Slatkin1987}), there may have been too few sites on most islands to conclusively detect it, and even then, the absence of detectable isolation-by-distance may not necessarily constitute evidence for the absence of drift. Besides, spatial autocorrelation was the strongest on islands sampled at the largest scales (e.g. some sites on Abaco were nearly 100km apart, and Eleuthera -- the second strongest signal, albeit nonsignificant -- had sites more than 30km apart), such that it is possible that neutral processes and/or dispersal limitations might contribute to shaping variation over long distances. That said, many significant differences were found between habitats in close proximity, contrary to what would be expected under isolation-by-distance, including on islands where spatial autocorrelation was detected. Moreover, \textit{A. sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations particularly prone to drift. Together with the fact that isolation-by-distance may not necessarily only emerge from drift (e.g. if there is a spatial environmental gradient), this indicates that genetic drift may have limited potential to explain the differences observed between habitats, at least at a local scale.}\\

### 2021-12-17 ms/sections/introduction.tex
implemented Graham's comments on the pdf

Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km. Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. If detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.

>>>>

Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km. \textcolor{olive}{While this species has traditionally been considered territorial, recent study reveals that they are polygynandrous and that gene flow is not impeded by territorial-like behaviors exibited by some males (Kamath and Losos 2018).} Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. If detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.

### 2021-12-17 ms/sections/results.tex
implemented Graham's comments on the pdf

A striking feature of our data was inconsistencies in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction. For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig. \ref{fig:CaymanBrac}D), on Abaco dewlaps from beach scrub and primary coppice were often mistaken, and on Bimini beach scrub dewlaps were more often classified into primary coppice or mangrove than into beach srub (Fig. \ref{fig:Bimini}D). In terms of variable importance, for multiple islands the random forests used information in the UV range to discriminate between at least some habitats, particularly on Abaco (Fig. \ref{fig:Abaco_supplement}F), Bimini (Fig. \ref{fig:Bimini}F), Cayman Brac (Fig. \ref{fig:CaymanBrac}F), Little Cayman (Fig. \ref{fig:LittleCayman}F) and Long Island (Fig. \ref{fig:LongIsland}F), but differences in UV reflectance involved different habitats and were in different directions among these islands.\\

>>>>

A striking feature of our data was inconsistenc\sout{ies}\textcolor{olive}{y} in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction. For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig. \ref{fig:CaymanBrac}D), on Abaco dewlaps from beach scrub and primary coppice were often mistaken, and on Bimini beach scrub dewlaps were more often classified into primary coppice or mangrove than into beach srub (Fig. \ref{fig:Bimini}D). In terms of variable importance, for multiple islands the random forests used information in the UV range to discriminate between at least some habitats, particularly on Abaco (Fig. \ref{fig:Abaco_supplement}F), Bimini (Fig. \ref{fig:Bimini}F), Cayman Brac (Fig. \ref{fig:CaymanBrac}F), Little Cayman (Fig. \ref{fig:LittleCayman}F) and Long Island (Fig. \ref{fig:LongIsland}F), but differences in UV reflectance involved different habitats and were in different directions among these islands.\\
